\thetitle

### 2023-06-03 lang-design.tex
Save.

\noindent
and with the induction hypothesis
\lstinline[basicstyle=\fontsize{8}{9}\ttfamily]!\forall G', includedin (extend G x T1) G' -> hasty G' t T2!.
%
The programmer can use tactic programming to discharge the goal.
%The idiomatic programming experience with Coq is largely preserved.
Because the lemma is by \lsti!FInduction! on "hasty" and because "STLCFix" adds
a constructor "ht_fix" to "hasty", the programmer is required in family "STLCFix" to
extend the proof of "substlem" to handle this extra case.

@@

As analyzed earlier, ordinary recursors, such as "tm_rect", are impossible
within a family in which the name of the inductive type is late bound.
However,
a key observation is that extensible inductive types still admit a
weakened elimination principle where the motive is an \lsti{option} type.
For example, within family "STLC", the partial recursor for
"tm" has the following type:

\begin{centered}
\begin{minipage}{\textwidth}
\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
  [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
  [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] -> \forall t, option (P t)
\end{lstlisting}
\end{minipage}
\end{centered}

@@

Partial recursor is simply adding partiality into the signature of the eliminator, by wrapping each motive with "option". This partial recursor seems non-canonical compared to the (dependent) eliminator for Inductive type as the latter can be considered as universal property. This extensionality makes dependent eliminator one of \textit{the most powerful} reasoning tools around inductive type.

Even though we cannot prove partial recursors are \textit{the
most powerful} reasoning tool for extensible inductive types,
the fact that propositional partial
recursors can derive injectivity alreadt demonstrate itself as a good \textit{extensional characterization} of
(non-indexed) extensible inductive type (under certain
constraints)---because we can ``embed'' vanilla inductive types into
types supporting a partial recursor: a bit more formally, and
restricting our focus to non-indexed inductive types:

>>>>

\noindent
and with the induction hypothesis
\lstinline[basicstyle=\fontsize{8}{9}\ttfamily]!\forall G', includedin (extend G x T1) G' -> hasty G' t T2!.
%
The programmer can use tactic programming to discharge the goal.
%The idiomatic programming experience with Coq is largely preserved.
Because the lemma is by \lsti!FInduction! on "hasty" and because "STLCFix" adds
a constructor "ht_fix" to "hasty", the programmer is required in family "STLCFix" to
extend the proof of "weakenlem" to handle this extra case.

@@

As analyzed earlier, ordinary recursors, such as "tm_rect", are impossible
within a family in which the name of the inductive type is late bound.
However,
a key observation is that extensible inductive types still admit a
weakened elimination principle where the motive is an \lsti{option} type.
For example, within family "STLC", the partial recursor for
"tm" has the following type, with extra equations describing
its behvaiour on each constuctors:

\begin{centered}
\begin{minipage}{\textwidth}
\newcommand\hlcase[1]{\tikzmarkin[disable rounded corners=true,set fill color=yellow!18,set border color=yellow!18]{#1}(0.00,-0.05)(-0.00,0.20)}
\begin{lstlisting}[basicstyle=\fontsize{8.25}{9.9}\ttfamily]
tm_prect_STLC : \forall (P : tm -> Type), [@\hlcase{tm-partialrec-case-unit}@]option (P tm_unit)[@\tikzmarkend{tm-partialrec-case-unit}@] -> [@\hlcase{tm-partialrec-case-var}@](\forall x, option (P (tm_var x)))[@\tikzmarkend{tm-partialrec-case-var}@] ->
  [@\hlcase{tm-partialrec-case-abs}@](\forall x t, option (P t) -> option (P (tm_abs x t)))[@\tikzmarkend{tm-partialrec-case-abs}@] ->
  [@\hlcase{tm-partialrec-case-app}@](\forall t1, option (P t1) -> \forall t2, option (P t2) -> option (P (tm_app t1 t2)))[@\tikzmarkend{tm-partialrec-case-app}@] -> \forall t, option (P t)
(* For arbitrary [@P@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@], and we denote tmPR := tm_prect_STLC [@P@] [@$r_1$@] [@$r_2$@] [@$r_3$@] [@$r_4$@] *)
(* We will have following equations *)
tm_prect_STLC_unit : tmPR tm_unit      = [@$r_1$@]
tm_prect_STLC_var  : tmPR (tm_var x)   = [@$r_2$@] x
tm_prect_STLC_abs  : tmPR (tm_abs x b) = [@$r_3$@] a b (tmPR b)
tm_prect_STLC_app  : tmPR (tm_app a b) = [@$r_4$@] a (tmPR a) b (tmPR b)
\end{lstlisting}
\end{minipage}
\end{centered}

This corresponds to the dependent eliminator for the non-extensible version of "tm".

@@

A partial recursor is simply adding partiality into the signature of the ordinary eliminator---by wrapping each motive with "option"---together 
with the equations describing its computational behaviors inherited from the ordinary eliminators.

This partial recursor seems non-canonical compared to the (dependent) eliminator for Inductive type as the latter can be considered as universal property. This extensionality makes dependent eliminator one of \textit{the most powerful} reasoning tools around inductive type.

Even though we cannot prove partial recursors are \textit{the
most powerful} reasoning tool for extensible inductive types,
the fact that propositional partial
recursors can derive injectivity already demonstrate itself as a good \textit{extensional characterization} of
(non-indexed) extensible inductive type (under certain
constraints)---because we can ``embed'' vanilla inductive types into
types supporting a partial recursor: a bit more formally, and
restricting our focus to non-indexed inductive types:
